Reconstructing the last known movements of one of Nature's giants

### 2018-10-12 manuscript/NatureComms/blue-whale-ms-oct18.tex
Fix abstract

\section{Abstract}
Understanding animal movements is crucial for effective conservation, especially as species ranges shift due to climate change\cite{runge2014conserving,robinson2009travelling}. 
Animal tissues housed in museum collections provide biochemical records of behaviour that can be used to reconstruct individual-level movements over long timescales and under historical climatic and ecological conditions\cite{newsome2010using}, but interpreting these records can be challenging. 
Here we combine stable isotope analyses of baleen with novel agent-based simulation models, to reconstruct seven years of movements of an iconic female blue whale (\textit{Balaenoptera musculus}) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London. 
We provide the longest continuous record of movement for an individual blue whale in the north Atlantic. 
We show that the whale spent at least two years resident in sub-tropical waters as a juvenile, before commencing annual latitudinal migrations consistent with patterns of movements of blue whales in the northeast Atlantic inferred from observations, tagging and historic whaling data. 
The isotopic record is consistent with observations of opportunistic feeding throughout the year in blue whales, and suggests pregnancy and calf-rearing in the year before death.
Historic tissue collections can provide unique insight into individual-level movement behaviours associated with changing climate, ecological conditions and anthropogenic pressures, and reveal aspects of life history. 
Combining biochemical tracers with simulation modeling unlocks the information recorded in these irreplaceable archives.

>>>>

\section{Abstract}
The spatial ecology of rare, migratory oceanic mammals such as the blue whale (\textit{Balaenoptera musculus}) is difficult to study directly, leading to distinct knowledge gaps associated with migratory behavior at the individual scale. 
Recent advances in satellite telemetry have improved our understanding of individual-level movement, but tags rarely provide data for more than a few months, and tagging cannot provide retrospective information. 
Incrementally-grown tissues provide alternative, biochemical records of behaviour that can be used to reconstruct individual-level movements over long timescales, but inferring detailed spatio-temporal information from biochemical tracers is challenging. 
Here we present a new approach combining stable isotope analyses of incrementally-grown tissues with simulation models to infer individual-level animal movements.  
We use our new method to reconstruct multiple years of movement histories of individual modern and historic blue whales, including an iconic blue whale that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London.
We recover inferences of individual space use and between-individual variation in movements that are consistent with satellite-tag based data for modern blue whales in the northeast Pacific Ocean. 
For the historic whale, inferred movements over seven years of life of a single individual are also consistent with our limited understanding of the movement ecology of blue whales in the northeast Atlantic Ocean. 
Combining biochemical tracers in incrementally-grown tissues with simulation modeling unlocks the information recorded in historic collections, and provides inferences that are complimentary to and comparable with tag-derived movement data.
